---
title: "2025treetemp_datawrangling"
author: "Erika Lee"
date: "2025-02-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lterdatasampler)
library(dplyr)
library(ggplot2)
library(readr)
library(readxl)
library(lubridate)
library(plotly)
library(openxlsx)
library(plotly)
library(rstatix)
library(htmlwidgets)
library(RColorBrewer)
library(patchwork)
library(ggpubr)
library(suncalc)

setwd("/Volumes/wcnr-network/Research/Kampf/Private/field_data")
```

```{r}
#to clear prior environment
rm(list = ls())
```

# Calling in Data

\*\* bs means below snow -\> all new 1 and 0.5 m sensors!

```{r}
#calling in tree temp data from n-drive
## burned
bs_pers_b_alldata <- read_csv("nsf/105E_pers_burned/105E_pers_b_bs_composite.csv")

bs_trans_b_alldata <- read_csv("nsf/105E_trans_burned/105E_trans_b_bs_composite.csv")

bs_pers_ub_alldata <- read_csv("nsf/105E_pers_unburned/105E_pers_ub_bs_composite.csv")

## don't have yet!
#bs_trans_ub_alldata <- read_csv("nsf/105E_trans_unburned/105E_trans_ub__bs_composite.csv")
```

2m sensor data

```{r}
#read in temperature data
pers_b_tree_composite=read_excel("nsf/105E_pers_burned/105E_pers_b_composite.xlsx")

#filter temp data to full data frame
pers_b_tree_fulldata <- pers_b_tree_composite %>%
  mutate(datetime = ymd_hms(TIMESTAMP, tz = "MST")) %>%
  select(datetime, everything()) %>%
  select(-c(TIMESTAMP, RECORD))

#adding in an aspect column - WITH ONLY 14 SENSORS!
pers_b_aspects <- pers_b_tree_fulldata %>%
  pivot_longer(cols = starts_with("Temp"),
               names_to = "sensor_id",
               values_to = "temp") %>%
  #add in aspect column
  mutate(aspect = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(1)",
      "Temp_C_Avg(5)",
      "Temp_C_Avg(9)",
      "Temp_C_Avg(13)", "Temp_C_Avg(15)"
    ),
    "north",
    "other"
  )) %>%
  mutate(aspect = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(2)",
      "Temp_C_Avg(6)",
      "Temp_C_Avg(10)",
      "Temp_C_Avg(14)", "Temp_C_Avg(16)"),
    "south",
    aspect
  )) %>%
  mutate(aspect = ifelse(
    sensor_id %in% c("Temp_C_Avg(3)", "Temp_C_Avg(7)", "Temp_C_Avg(11)"),
    "charred",
    aspect
  )) %>%
  mutate(aspect = ifelse(
    sensor_id %in% c("Temp_C_Avg(4)", "Temp_C_Avg(8)", "Temp_C_Avg(12)"),
    "uncharred",
    aspect
  )) %>%
  #adding in sensor name column
  mutate(tree_name = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(1)",
      "Temp_C_Avg(2)",
      "Temp_C_Avg(3)",
      "Temp_C_Avg(4)"
    ),
    "gb_w",
    "other"
  )) %>%
  mutate(tree_name = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(5)",
      "Temp_C_Avg(6)",
      "Temp_C_Avg(7)",
      "Temp_C_Avg(8)"
    ),
    "gb_e",
    tree_name
  )) %>%
  mutate(tree_name = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(9)",
      "Temp_C_Avg(10)",
      "Temp_C_Avg(11)",
      "Temp_C_Avg(12)"
    ),
    "gb_s",
    tree_name
  )) %>%
  mutate(tree_name = ifelse(
    sensor_id %in% c("Temp_C_Avg(13)", "Temp_C_Avg(14)"),
    "db_n",
    tree_name
  )) %>%
  mutate(tree_name = ifelse(
    sensor_id %in% c("Temp_C_Avg(15)", "Temp_C_Avg(16)"),
    "db_s",
    tree_name
  )) %>%
#adding in sensor height & zone
  mutate(sensor_height_m = 2.0, zone = "PSZ", day = as.Date(datetime, tz = "MST")) %>%
  #removing charred/uncharred sensors
  filter(!aspect %in% c("charred", "uncharred")) %>%
    #adding in burn status column
  mutate(burn_status = case_when(
    str_starts(tree_name, "gb_") ~ "GB",
    str_starts(tree_name, "db_") ~ "DB",
    TRUE ~ NA_character_  # Or any default value you want
  )) %>%
  rename(temp_C = "temp") %>%
  select(-sensor_id) %>%
  #filtering timeperiod to match BS dataframe 
  filter(datetime >= "2024-12-18 14:30:00") %>%
  filter(!tree_name == "gb_w") %>%
  #adding phase column 
  mutate(day = as.Date(day, format = ("%Y-%m-%d"))) %>%
  inner_join(p_b_solar_times_2025_V2, by = "day") %>%
  #creating a phase column that parses by daily sunrise/sunset times
  mutate(
    phase = case_when(
      datetime >= sunrise & datetime < sunset ~ "day",
      datetime < sunrise ~ "night",
      datetime >= sunset ~ "night",
      TRUE ~ NA_character_)) %>%
  select(datetime, day, phase, zone, burn_status, tree_name, aspect, sensor_height_m, temp_C, solarNoon, sunrise, sunset, dawn, dusk)
```

```{r}
#read in temperature data
trans_b_tree_composite=read_excel("nsf/105E_trans_burned/105E_trans_b_composite.xlsx")

#filter temp data to full data frame
trans_b_tree_fulldata <- trans_b_tree_composite %>%
  mutate(datetime = ymd_hms(TIMESTAMP, tz = "MST")) %>%
  select(datetime, everything()) %>%
  select(-c(TIMESTAMP, Record)) %>%
  na.omit()

#adding in an aspect column - ** need to edit date on trans_b sensors, seems weird, should be 2023.12.06 around 10 am?
trans_b_aspects <- trans_b_tree_fulldata %>%
  pivot_longer(cols = starts_with("Temp"),
               names_to = "sensor_id",
               values_to = "temp") %>%
  #add in aspect column
  mutate(aspect = ifelse(
    sensor_id %in% c("Temp_C_Avg(1)", "Temp_C_Avg(5)", "Temp_C_Avg(9)", "Temp_C_Avg(11)"),
    "north",
    "other"
  )) %>%
mutate(aspect = ifelse(
    sensor_id %in% c("Temp_C_Avg(2)", "Temp_C_Avg(6)", "Temp_C_Avg(10)", "Temp_C_Avg(12)"),
    "south",
    aspect)) %>%
  mutate(aspect = ifelse(
    sensor_id %in% c("Temp_C_Avg(3)", "Temp_C_Avg(7)"),
    "charred",
    aspect)) %>%
    mutate(aspect = ifelse(
    sensor_id %in% c("Temp_C_Avg(4)", "Temp_C_Avg(8)"),
    "uncharred",
    aspect)) %>%
  #adding in sensor name column
  mutate(tree_name = ifelse(
    sensor_id %in% c("Temp_C_Avg(1)", "Temp_C_Avg(2)", "Temp_C_Avg(3)", "Temp_C_Avg(4)"),
    "gb_n",
    "other"
  )) %>%
  mutate(tree_name = ifelse(
    sensor_id %in% c("Temp_C_Avg(5)", "Temp_C_Avg(6)", "Temp_C_Avg(7)", "Temp_C_Avg(8)"),
    "gb_s",
    tree_name)) %>%
  mutate(tree_name = ifelse(
    sensor_id %in% c("Temp_C_Avg(9)", "Temp_C_Avg(10)"),
    "db_n",
    tree_name)) %>%
    mutate(tree_name = ifelse(
    sensor_id %in% c("Temp_C_Avg(11)", "Temp_C_Avg(12)"),
    "db_s",
    tree_name)) %>%
  #adding in sensor height, zone & date
  mutate(sensor_height_m = 2.0, zone = "TSZ", day = as.Date(datetime, tz = "MST")) %>%
  #removing charred/uncharred sensors
  filter(!aspect %in% c("charred", "uncharred")) %>%
  #adding in burn status column
  mutate(burn_status = case_when(
    str_starts(tree_name, "gb_") ~ "GB",
    str_starts(tree_name, "db_") ~ "DB",
    TRUE ~ NA_character_  # Or any default value you want
  )) %>%
  rename(temp_C = "temp") %>%
  select(-sensor_id) %>%
  #filtering date so that it matches bs dataset
  filter(datetime >= as.POSIXct("2024-11-13 11:30:00", tz = "MST")) %>%
  #adding phase column 
  mutate(day = as.Date(day, format = ("%Y-%m-%d"))) %>%
  inner_join(p_b_solar_times_2025_V2, by = "day") %>%
  #creating a phase column that parses by daily sunrise/sunset times
  mutate(
    phase = case_when(
      datetime >= sunrise & datetime < sunset ~ "day",
      datetime < sunrise ~ "night",
      datetime >= sunset ~ "night",
      TRUE ~ NA_character_)) %>%
  select(datetime, day, phase, zone, burn_status, tree_name, aspect, sensor_height_m, temp_C, solarNoon, sunrise, sunset, dawn, dusk)
```

# Data Wrangling

Instructions: Call in this data & rerun all the steps whenever I add new data to the composite files in nsf field data drive.

-   need to change dates on X_solar_times data wrangling steps then rerun when I add data to the composite files -\> update time range to current range in composite csv's

-   need to export new \_r versions when I add data to composite csv's

```{r}
bs_pers_b_aspects <- bs_pers_b_alldata %>%
  mutate(TIMESTAMP = paste0(TIMESTAMP, ":00")) %>%  # Add missing seconds
  mutate(datetime = mdy_hms(TIMESTAMP, tz = "MST")) %>%
  select(-TIMESTAMP) %>%
  select(datetime, everything()) %>%
  pivot_longer(cols = matches("^GB|^DB"),  # Use matches() for multiple prefixes
  names_to = "tree_id",
  values_to = "temp_C") %>%
  #add in aspect column
  mutate(aspect = ifelse(
    tree_id %in% c(
      "GB_E_N_1_Avg",
      "GB_E_N_05_Avg",
      "DB_N_N_1_Avg",
      "DB_N_N_05_Avg"),
    "north",
    "south"
  )) %>%
  # Add tree_name column based on the start of tree_id
  mutate(tree_name = case_when(
    str_starts(tree_id, "GB_E") ~ "gb_e",  # If tree_id starts with "GB_E"
    str_starts(tree_id, "DB_N") ~ "db_n",  # If tree_id starts with "DB_N"
    TRUE ~ NA_character_  # If neither condition is met, set to NA
  )) %>%
  #add burn status column
   # Add tree_name column based on the start of tree_id
  mutate(burn_status = case_when(
    str_starts(tree_id, "GB_E") ~ "GB",  # If tree_id starts with "GB_E"
    str_starts(tree_id, "DB") ~ "DB",  # If tree_id starts with "DB_N"
    TRUE ~ NA_character_  # If neither condition is met, set to NA
  )) %>%
  mutate(sensor_height_m = case_when(
    str_ends(tree_id, "1_Avg") ~ 1,  # If tree_id ends with "1_Avg"
    str_ends(tree_id, "05_Avg") ~ 0.5,  # If tree_id ends with "05_Avg"
    TRUE ~ NA_real_  # If neither condition is met, set to NA
  )) %>%
  mutate(zone = "PSZ", day = as.Date(datetime, tz = "MST")) %>%
  select(datetime, day, zone, burn_status, tree_name, aspect, sensor_height_m, temp_C)
```

```{r}
bs_trans_b_aspects <- bs_trans_b_alldata %>%
  mutate(TIMESTAMP = paste0(TIMESTAMP, ":00")) %>%  # Add missing seconds
  mutate(datetime = mdy_hms(TIMESTAMP, tz = "MST")) %>%
  select(-TIMESTAMP) %>%
  select(datetime, everything()) %>%
  pivot_longer(cols = matches("^GB|^DB"),  # Use matches() for multiple prefixes
  names_to = "tree_id",
  values_to = "temp_C") %>%
  #add in aspect column
  mutate(aspect = ifelse(
    tree_id %in% c(
      "gb_s_n_1_Avg",
      "gb_s_n_05_Avg",
      "db_s_n_1_Avg",
      "db_s_n_05_Avg"),
    "north",
    "south"
  )) %>%
  # Add tree_name column based on the start of tree_id
  mutate(tree_name = case_when(
    str_starts(tree_id, "gb_s") ~ "gb_s",  # If tree_id starts with "GB_E"
    str_starts(tree_id, "db_s") ~ "db_s",  # If tree_id starts with "DB_N"
    TRUE ~ NA_character_  # If neither condition is met, set to NA
  )) %>%
  #add burn status column
   # Add tree_name column based on the start of tree_id
  mutate(burn_status = case_when(
    str_starts(tree_id, "gb") ~ "GB",  # If tree_id starts with "GB_E"
    str_starts(tree_id, "db") ~ "DB",  # If tree_id starts with "DB_N"
    TRUE ~ NA_character_  # If neither condition is met, set to NA
  )) %>%
  mutate(sensor_height_m = case_when(
    str_ends(tree_id, "1_Avg") ~ 1,  # If tree_id ends with "1_Avg"
    str_ends(tree_id, "05_Avg") ~ 0.5,  # If tree_id ends with "05_Avg"
    TRUE ~ NA_real_  # If neither condition is met, set to NA
  )) %>%
  mutate(zone = "TSZ", day = as.Date(datetime, tz = "MST")) %>%
  select(datetime, day, zone, burn_status, tree_name, aspect, sensor_height_m, temp_C)
```

```{r}
bs_pers_ub_aspects <- bs_pers_ub_alldata %>%
  mutate(TIMESTAMP = paste0(TIMESTAMP, ":00")) %>%  # Add missing seconds
  mutate(datetime = mdy_hms(TIMESTAMP, tz = "MST")) %>%
  select(-TIMESTAMP) %>%
  select(datetime, everything()) %>%
  pivot_longer(cols = matches("^UB"),  # Use matches() for multiple prefixes
  names_to = "tree_id",
  values_to = "temp_C") %>%
  #add in aspect column
  mutate(aspect = ifelse(
    tree_id %in% c(
      "UB_N_N_2_Avg",
      "UB_N_N_05_Avg"), 
    "north",
    "south"
  )) %>%
  # Add tree_name column based on the start of tree_id
  mutate(tree_name = "ub_n", burn_status = "live", zone = "PSZ") %>%
  mutate(sensor_height_m = case_when(
    str_ends(tree_id, "2_Avg") ~ 2,  # If tree_id ends with "1_Avg"
    str_ends(tree_id, "05_Avg") ~ 0.5,  # If tree_id ends with "05_Avg"
    TRUE ~ NA_real_  # If neither condition is met, set to NA
  )) %>%
  mutate(day = as.Date(datetime, tz = "MST")) %>%
  select(datetime, day, zone, burn_status, tree_name, aspect, sensor_height_m, temp_C)
```

```{r}
## WILL NEED TO EDIT THIS ONCE I HAVE DATA FROM SITE!
bs_trans_ub_aspects <- bs_trans_ub_alldata %>%
  mutate(TIMESTAMP = paste0(TIMESTAMP, ":00")) %>%  # Add missing seconds
  mutate(datetime = mdy_hms(TIMESTAMP, tz = "MST")) %>%
  select(-TIMESTAMP) %>%
  select(datetime, everything()) %>%
  pivot_longer(cols = matches("^UB"),  # Use matches() for multiple prefixes
  names_to = "tree_id",
  values_to = "temp_C") %>%
  #add in aspect column
  mutate(aspect = ifelse(
    tree_id %in% c(
      "UB_N_N_2_Avg",
      "UB_N_N_05_Avg"), 
    "north",
    "south"
  )) %>%
  # Add tree_name column based on the start of tree_id
  mutate(tree_name = "ub_n", burn_status = "live", zone = "TSZ") %>%
  mutate(sensor_height_m = case_when(
    str_ends(tree_id, "2_Avg") ~ 2,  # If tree_id ends with "1_Avg"
    str_ends(tree_id, "05_Avg") ~ 0.5,  # If tree_id ends with "05_Avg"
    TRUE ~ NA_real_  # If neither condition is met, set to NA
  )) %>%
  mutate(day = as.Date(datetime, tz = "MST")) %>%
  select(datetime, day, zone, burn_status, tree_name, aspect, sensor_height_m, temp_C)
```

\*\* add in phase next! Need to download p_solar_times then do a mutate to add phase to all of the above datasets -\> but make a new version labeled XXX_r?

Adding in phase

```{r}
# Specify PSZ and TSZ locations
##pers_b - from 105E location!
p_b_lat <- 40.56379284
p_b_lon <- -105.867215
##pers_ub - from 105E locations
p_ub_lat <- 40.5638075
p_ub_lon <- -105.8705715

##trans_b - from 105E locations!
t_b_lat <- 40.60600629
t_b_lon <- -105.8333335

##trans_ub - from 105E locations!
t_ub_lat <- 40.568396
t_ub_lon <- -105.587383

## use the p_b_lat/long since all of these dates/times are almost identical for sunrise/sunset

# Create a date sequence from February 2nd to May 24th, 2024 for solar exposure
solex_dates <- seq(as.Date("2024-10-01"), as.Date("2025-07-01"), by = "day")

# Get solar times for each day
p_b_solar_times_2025 <- getSunlightTimes(date = solex_dates, lat = p_b_lat, lon = p_b_lon, tz = "MST") %>%
  mutate(
    day = as.Date(date),  # Ensure 'day' is of Date type for joining
    sunrise = with_tz(as.POSIXct(sunrise), tzone = "MST"),  # Set MST timezone
    sunset = with_tz(as.POSIXct(sunset), tzone = "MST")     # Set MST timezone
  )
```

```{r}
#cutting the sunrise/sunset times to the current date range
#PSZ burned
p_b_solar_times_2025_V2 <- p_b_solar_times_2025 %>%
  mutate(day = as.Date(date)) %>%
  filter(day >= as.Date("2024-12-18") & day <= as.Date("2025-06-30")) %>% 
  mutate(day = as.Date(day, format = "%Y-%m-%d"))

#TSZ burned
t_b_solar_times_2025 <- p_b_solar_times_2025 %>%
  mutate(day = as.Date(date)) %>%
  filter(day >= as.Date("2024-11-13") & day <= as.Date("2025-06-30")) %>% 
  mutate(day = as.Date(day, format = "%Y-%m-%d"))

#PSZ unburned
p_ub_solar_times_2025 <- p_b_solar_times_2025 %>%
  mutate(day = as.Date(date)) %>%
  filter(day >= as.Date("2024-11-12") & day <= as.Date("2025-06-30")) %>% 
  mutate(day = as.Date(day, format = "%Y-%m-%d"))

#TSZ unburned - DONT HAVE DATA YET!
t_ub_solar_times_2025 <- p_b_solar_times_2025 %>%
  mutate(day = as.Date(date)) %>%
  filter(day >= as.Date("2024-11-12") & day <= as.Date("2025-06-30")) %>% 
  mutate(day = as.Date(day, format = "%Y-%m-%d"))
```

```{r}
#PSZ burned data with day/night
bs_pers_b_aspects_filtered <- bs_pers_b_aspects %>%
  mutate(day = as.Date(day, format = ("%Y-%m-%d"))) %>%
  inner_join(p_b_solar_times_2025_V2, by = "day") %>%
  #creating a phase column that parses by daily sunrise/sunset times
  mutate(
    phase = case_when(
      datetime >= sunrise & datetime < sunset ~ "day",
      datetime < sunrise ~ "night",
      datetime >= sunset ~ "night",
      TRUE ~ NA_character_)) %>%
  select(datetime, day, phase, zone, burn_status, tree_name, aspect, sensor_height_m, temp_C, solarNoon, sunrise, sunset, dawn, dusk)
```

```{r}
#TSZ burned data with day/night
bs_trans_b_aspects_filtered <- bs_trans_b_aspects %>%
  mutate(day = as.Date(day, format = ("%Y-%m-%d"))) %>%
  inner_join(t_b_solar_times_2025, by = "day") %>%
  #creating a phase column that parses by daily sunrise/sunset times
  mutate(
    phase = case_when(
      datetime >= sunrise & datetime < sunset ~ "day",
      datetime < sunrise ~ "night",
      datetime >= sunset ~ "night",
      TRUE ~ NA_character_)) %>%
  select(datetime, day, phase, zone, burn_status, tree_name, aspect, sensor_height_m, temp_C, solarNoon, sunrise, sunset, dawn, dusk) %>%
  filter(datetime >= as.POSIXct("2024-11-13 11:30:00", tz = "MST"))
```

```{r}
#PSZ unburned data with day/night
bs_pers_ub_aspects_filtered <- bs_pers_ub_aspects %>%
  mutate(day = as.Date(day, format = ("%Y-%m-%d"))) %>%
  inner_join(p_ub_solar_times_2025, by = "day") %>%
  #creating a phase column that parses by daily sunrise/sunset times
  mutate(
    phase = case_when(
      datetime >= sunrise & datetime < sunset ~ "day",
      datetime < sunrise ~ "night",
      datetime >= sunset ~ "night",
      TRUE ~ NA_character_)) %>%
  select(datetime, day, phase, zone, burn_status, tree_name, aspect, sensor_height_m, temp_C, solarNoon, sunrise, sunset, dawn, dusk)
```

Exporting filtered datasets

```{r}
#datasets with full extent as of now, but will need to rerun when I upload new data to the composite files
write_csv(bs_pers_b_aspects_filtered, "nsf/105E_pers_burned/105E_pers_b_bs_r.csv")

write_csv(bs_trans_b_aspects_filtered, "nsf/105E_trans_burned/105E_trans_b_bs_r.csv")

write_csv(bs_pers_ub_aspects_filtered, "nsf/105E_pers_unburned/105E_pers_ub_bs_r.csv")

write_csv(bs_trans_ub_aspects_filtered, "nsf/105E_trans_unburned/105E_trans_ub_bs_r.csv")
```

Joining bs and 2m data for full dataset

```{r}
#PSZ burned
pers_full_treetemp_data_2025 <- bs_pers_b_aspects_filtered %>%
  full_join(pers_b_aspects, by = c("datetime", "day", "zone", "burn_status", "tree_name", "aspect", "phase", "sensor_height_m", "temp_C", "solarNoon", "sunrise", "sunset", "dawn", "dusk")) %>%
  full_join(bs_pers_ub_aspects_filtered, by = c("datetime", "day", "zone", "burn_status", "tree_name", "aspect", "phase", "sensor_height_m", "temp_C", "solarNoon", "sunrise", "sunset", "dawn", "dusk"))
```

```{r}
trans_full_treetemp_data_2025 <- bs_trans_b_aspects_filtered %>%
  full_join(trans_b_aspects, by = c("datetime", "day", "zone", "burn_status", "tree_name", "aspect", "phase", "sensor_height_m", "temp_C", "solarNoon", "sunrise", "sunset", "dawn", "dusk"))

#adding trans UB - cant do this yet because I don't have data!
full_join(bs_trans_ub_aspects_filtered, by = c("datetime", "day", "zone", "burn_status", "tree_name", "aspect", "phase", "sensor_height_m", "temp_C", "solarNoon", "sunrise", "sunset", "dawn", "dusk"))
```

# Data Exploration

## Three Day Comparison

TSZ

```{r}
TSZ_threeday_comp <- trans_full_treetemp_data_2025 %>%
  filter(datetime >= "2025-03-18 00:00:00" & datetime <= "2025-03-21 00:00:00") %>%
  mutate(sensor_height_m = as.character(sensor_height_m)) %>%
  mutate(sensor_height_m = factor(sensor_height_m, levels = c("2", "1", "0.5")))
```

```{r}
#three day lineplot PSZ
TSZ_threeday_comp_lineplot <- ggplot(data = TSZ_threeday_comp) + 
        # Bold line at y = 0
  geom_hline(yintercept = 0, linetype = "solid", color = "grey23", size = 0.3) +
  geom_line(aes(x = datetime, y = temp_C, linetype = sensor_height_m, color = burn_status, alpha = sensor_height_m), linewidth = 0.7) +
  facet_wrap(~aspect) +
  scale_linetype_manual(
    name = "Sensor Height (m)",  # New legend title (showing as 'Tree' instead of 'Tree Name')
    values = c("2" = "solid", "1" = "dashed", "0.5" = "dotted"),
    labels = c("2", "1", "0.5")  # Ensuring the legend shows the correct labels
  ) + 
  scale_color_manual(
    name = "Burn Status",
    values = c("GB" = "#1B7837", "DB" = "#D95F02", "live" = alpha("#2166AC", 0.8))) +
  scale_alpha_manual(
  values = c("2" = 0.7, "1" = 1, "0.5" = 1),  # Make 2m more transparent
  guide = "none") +
  theme_bw() +
  labs(title = "b) TSZ", x = NULL, y = "Temperature (째C)") +
  theme(
    strip.text = element_text(size = 11),  # Facet title size
    axis.title = element_text(size = 11),  # Axis title size
    axis.text = element_text(size = 11),  # Axis tick labels
    legend.title = element_text(size = 11),  # Legend title size
    legend.text = element_text(size = 10)  # Legend labels size
  )
 
TSZ_threeday_comp_lineplot
```

PSZ

\*\*not working well right now because I don't have a consistent dataset.

```{r}
PSZ_threeday_comp <- pers_full_treetemp_data_2025 %>%
  filter(datetime >= "2025-04-11 00:00:00" & datetime <= "2025-04-14 00:00:00") %>%
  mutate(sensor_height_m = as.character(sensor_height_m)) %>%
  mutate(sensor_height_m = factor(sensor_height_m, levels = c("2", "1", "0.5")))
```

```{r}
#three day lineplot PSZ
PSZ_threeday_comp_lineplot <- ggplot(data = PSZ_threeday_comp) + 
        # Bold line at y = 0
  geom_hline(yintercept = 0, linetype = "solid", color = "grey23", size = 0.3) +
  geom_line(aes(x = datetime, y = temp_C, linetype = sensor_height_m, color = burn_status, alpha = sensor_height_m), linewidth = 0.7) +
  facet_wrap(~aspect) +
  scale_linetype_manual(
    name = "Sensor Height (m)",  # New legend title (showing as 'Tree' instead of 'Tree Name')
    values = c("2" = "solid", "1" = "dashed", "0.5" = "dotted"),
    labels = c("2", "1", "0.5")  # Ensuring the legend shows the correct labels
  ) + 
  scale_color_manual(
    name = "Burn Status",
    values = c("GB" = "#1B7837", "DB" = "#D95F02", "live" = alpha("#2166AC", 0.8))) +
  scale_alpha_manual(
  values = c("2" = 0.7, "1" = 1, "0.5" = 1),  # Make 2m more transparent
  guide = "none") +
  theme_bw() +
  labs(title = "a) PSZ", x = NULL, y = "Temperature (째C)") +
  theme(
    strip.text = element_text(size = 11),  # Facet title size
    axis.title = element_text(size = 11),  # Axis title size
    axis.text = element_text(size = 11),  # Axis tick labels
    legend.title = element_text(size = 11),  # Legend title size
    legend.text = element_text(size = 10)  # Legend labels size
  )
 
PSZ_threeday_comp_lineplot
```

Full Dataset

Mean day and night temps

```{r}
#daytime dataframe
TSZ_full_mean_daily_dt <- trans_full_treetemp_data_2025 %>%
  filter(phase == "day", datetime >= "2025-01-03 00:00:00" & datetime <= "2025-04-15 00:00:00") %>%
  mutate(sensor_height_m = as.character(sensor_height_m)) %>%
  mutate(sensor_height_m = factor(sensor_height_m, levels = c("2", "1", "0.5"))) %>%
  group_by(day, phase, zone, burn_status, aspect, sensor_height_m) %>%
  summarize(mean_dt_temp = mean(temp_C, na.rm = TRUE))
```

```{r}
#nighttime dataframe
TSZ_full_mean_daily_nt <- trans_full_treetemp_data_2025 %>%
  filter(phase == "night", datetime >= "2025-01-03 00:00:00" & datetime <= "2025-04-15 00:00:00") %>%
  mutate(sensor_height_m = as.character(sensor_height_m)) %>%
  mutate(sensor_height_m = factor(sensor_height_m, levels = c("2", "1", "0.5"))) %>%
  group_by(day, phase, zone, burn_status, aspect, sensor_height_m) %>%
  summarize(mean_dt_temp = mean(temp_C, na.rm = TRUE))
```

Plotting

```{r}
#daytime plot
TSZ_full_mean_dt_lineplot <- ggplot(data = TSZ_full_mean_daily_dt) + 
        # Bold line at y = 0
  geom_hline(yintercept = 0, linetype = "solid", color = "grey23", size = 0.3) +
  geom_line(aes(x = day, y = mean_dt_temp, linetype = sensor_height_m, color = burn_status, alpha = sensor_height_m), linewidth = 0.7) +
  facet_wrap(~aspect) +
  scale_linetype_manual(
    name = "Sensor Height (m)",  # New legend title (showing as 'Tree' instead of 'Tree Name')
    values = c("2" = "solid", "1" = "dashed", "0.5" = "dotted"),
    labels = c("2", "1", "0.5")  # Ensuring the legend shows the correct labels
  ) + 
  scale_color_manual(
    name = "Burn Status",
    values = c("GB" = "#1B7837", "DB" = "#D95F02", "live" = alpha("#2166AC", 0.8))) +
  scale_alpha_manual(
  values = c("2" = 0.7, "1" = 1, "0.5" = 1),  # Make 2m more transparent
  guide = "none") +
  theme_bw() +
  labs(title = "a) TSZ Daily Mean Day Temps", x = NULL, y = "Temperature (째C)") +
  theme(
    strip.text = element_text(size = 11),  # Facet title size
    axis.title = element_text(size = 11),  # Axis title size
    axis.text = element_text(size = 11),  # Axis tick labels
    legend.title = element_text(size = 11),  # Legend title size
    legend.text = element_text(size = 10)  # Legend labels size
  )
 
TSZ_full_mean_dt_lineplot
```

```{r}
#daytime plot
TSZ_full_mean_nt_lineplot <- ggplot(data = TSZ_full_mean_daily_nt) + 
        # Bold line at y = 0
  geom_hline(yintercept = 0, linetype = "solid", color = "grey23", size = 0.3) +
  geom_line(aes(x = day, y = mean_dt_temp, linetype = sensor_height_m, color = burn_status, alpha = sensor_height_m), linewidth = 0.7) +
  facet_wrap(~aspect) +
  scale_linetype_manual(
    name = "Sensor Height (m)",  # New legend title (showing as 'Tree' instead of 'Tree Name')
    values = c("2" = "solid", "1" = "dashed", "0.5" = "dotted"),
    labels = c("2", "1", "0.5")  # Ensuring the legend shows the correct labels
  ) + 
  scale_color_manual(
    name = "Burn Status",
    values = c("GB" = "#1B7837", "DB" = "#D95F02", "live" = alpha("#2166AC", 0.8))) +
  scale_alpha_manual(
  values = c("2" = 0.7, "1" = 1, "0.5" = 1),  # Make 2m more transparent
  guide = "none") +
  theme_bw() +
  labs(title = "b) TSZ Daily Mean Night Temps", x = NULL, y = "Temperature (째C)") +
  theme(
    strip.text = element_text(size = 11),  # Facet title size
    axis.title = element_text(size = 11),  # Axis title size
    axis.text = element_text(size = 11),  # Axis tick labels
    legend.title = element_text(size = 11),  # Legend title size
    legend.text = element_text(size = 10)  # Legend labels size
  )
 
TSZ_full_mean_nt_lineplot
```

```{r}
TSZ_comb_mean_dtnt_temps <- ggarrange(TSZ_full_mean_dt_lineplot, TSZ_full_mean_nt_lineplot, 
                                      nrow = 2, ncol = 1, common.legend = TRUE, legend = "right")

TSZ_comb_mean_dtnt_temps
```
