---
title: "2025treetemp_datawrangling"
author: "Erika Lee"
date: "2025-02-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lterdatasampler)
library(dplyr)
library(ggplot2)
library(readr)
library(readxl)
library(lubridate)
library(plotly)
library(openxlsx)
library(plotly)
library(rstatix)
library(htmlwidgets)
library(RColorBrewer)
library(patchwork)
library(ggpubr)
library(suncalc)

setwd("/Volumes/wcnr-network/Research/Kampf/Private/field_data")
```

```{r}
#to clear prior environment
rm(list = ls())
```

# Calling in Data

\*\* bs means below snow -\> all new 1 and 0.5 m sensors!

```{r}
#calling in tree temp data from n-drive
## burned
bs_pers_b_alldata <- read_csv("nsf/105E_pers_burned/105E_pers_b_bs_composite.csv")

bs_trans_b_alldata <- read_csv("nsf/105E_trans_burned/105E_trans_b_bs_composite.csv")

bs_pers_ub_alldata <- read_csv("nsf/105E_pers_unburned/105E_pers_ub_bs_composite.csv")

## don't have yet!
bs_trans_ub_alldata <- read_csv("nsf/105E_trans_unburned/105E_trans_ub__bs_composite.csv")
```

# Data Wrangling

Instructions: Call in this data & rerun all the steps whenever I add new data to the composite files in nsf field data drive.

-   need to change dates on X_solar_times data wrangling steps then rerun when I add data to the composite files -\> update time range to current range in composite csv's

-   need to export new \_r versions when I add data to composite csv's

```{r}
bs_pers_b_aspects <- bs_pers_b_alldata %>%
  mutate(TIMESTAMP = paste0(TIMESTAMP, ":00")) %>%  # Add missing seconds
  mutate(datetime = mdy_hms(TIMESTAMP, tz = "MST")) %>%
  select(-TIMESTAMP) %>%
  select(datetime, everything()) %>%
  pivot_longer(cols = matches("^GB|^DB"),  # Use matches() for multiple prefixes
  names_to = "tree_id",
  values_to = "temp_C") %>%
  #add in aspect column
  mutate(aspect = ifelse(
    tree_id %in% c(
      "GB_E_N_1_Avg",
      "GB_E_N_05_Avg",
      "DB_N_N_1_Avg",
      "DB_N_N_05_Avg"),
    "north",
    "south"
  )) %>%
  # Add tree_name column based on the start of tree_id
  mutate(tree_name = case_when(
    str_starts(tree_id, "GB_E") ~ "gb_e",  # If tree_id starts with "GB_E"
    str_starts(tree_id, "DB_N") ~ "db_n",  # If tree_id starts with "DB_N"
    TRUE ~ NA_character_  # If neither condition is met, set to NA
  )) %>%
  #add burn status column
   # Add tree_name column based on the start of tree_id
  mutate(burn_status = case_when(
    str_starts(tree_id, "GB_E") ~ "GB",  # If tree_id starts with "GB_E"
    str_starts(tree_id, "DB") ~ "DB",  # If tree_id starts with "DB_N"
    TRUE ~ NA_character_  # If neither condition is met, set to NA
  )) %>%
  mutate(sensor_height_m = case_when(
    str_ends(tree_id, "1_Avg") ~ 1,  # If tree_id ends with "1_Avg"
    str_ends(tree_id, "05_Avg") ~ 0.5,  # If tree_id ends with "05_Avg"
    TRUE ~ NA_real_  # If neither condition is met, set to NA
  )) %>%
  mutate(zone = "PSZ", day = as.Date(datetime, tz = "MST")) %>%
  select(datetime, day, zone, burn_status, tree_name, aspect, sensor_height_m, temp_C)
```

```{r}
bs_trans_b_aspects <- bs_trans_b_alldata %>%
  mutate(TIMESTAMP = paste0(TIMESTAMP, ":00")) %>%  # Add missing seconds
  mutate(datetime = mdy_hms(TIMESTAMP, tz = "MST")) %>%
  select(-TIMESTAMP) %>%
  select(datetime, everything()) %>%
  pivot_longer(cols = matches("^GB|^DB"),  # Use matches() for multiple prefixes
  names_to = "tree_id",
  values_to = "temp_C") %>%
  #add in aspect column
  mutate(aspect = ifelse(
    tree_id %in% c(
      "gb_s_n_1_Avg",
      "gb_s_n_05_Avg",
      "db_s_n_1_Avg",
      "db_s_n_05_Avg"),
    "north",
    "south"
  )) %>%
  # Add tree_name column based on the start of tree_id
  mutate(tree_name = case_when(
    str_starts(tree_id, "gb_s") ~ "gb_s",  # If tree_id starts with "GB_E"
    str_starts(tree_id, "db_s") ~ "db_s",  # If tree_id starts with "DB_N"
    TRUE ~ NA_character_  # If neither condition is met, set to NA
  )) %>%
  #add burn status column
   # Add tree_name column based on the start of tree_id
  mutate(burn_status = case_when(
    str_starts(tree_id, "gb") ~ "GB",  # If tree_id starts with "GB_E"
    str_starts(tree_id, "db") ~ "DB",  # If tree_id starts with "DB_N"
    TRUE ~ NA_character_  # If neither condition is met, set to NA
  )) %>%
  mutate(sensor_height_m = case_when(
    str_ends(tree_id, "1_Avg") ~ 1,  # If tree_id ends with "1_Avg"
    str_ends(tree_id, "05_Avg") ~ 0.5,  # If tree_id ends with "05_Avg"
    TRUE ~ NA_real_  # If neither condition is met, set to NA
  )) %>%
  mutate(zone = "TSZ", day = as.Date(datetime, tz = "MST")) %>%
  select(datetime, day, zone, burn_status, tree_name, aspect, sensor_height_m, temp_C)
```

```{r}
bs_pers_ub_aspects <- bs_pers_ub_alldata %>%
  mutate(TIMESTAMP = paste0(TIMESTAMP, ":00")) %>%  # Add missing seconds
  mutate(datetime = mdy_hms(TIMESTAMP, tz = "MST")) %>%
  select(-TIMESTAMP) %>%
  select(datetime, everything()) %>%
  pivot_longer(cols = matches("^UB"),  # Use matches() for multiple prefixes
  names_to = "tree_id",
  values_to = "temp_C") %>%
  #add in aspect column
  mutate(aspect = ifelse(
    tree_id %in% c(
      "UB_N_N_2_Avg",
      "UB_N_N_05_Avg"), 
    "north",
    "south"
  )) %>%
  # Add tree_name column based on the start of tree_id
  mutate(tree_name = "ub_n", burn_status = "live", zone = "PSZ") %>%
  mutate(sensor_height_m = case_when(
    str_ends(tree_id, "2_Avg") ~ 2,  # If tree_id ends with "1_Avg"
    str_ends(tree_id, "05_Avg") ~ 0.5,  # If tree_id ends with "05_Avg"
    TRUE ~ NA_real_  # If neither condition is met, set to NA
  )) %>%
  mutate(day = as.Date(datetime, tz = "MST")) %>%
  select(datetime, day, zone, burn_status, tree_name, aspect, sensor_height_m, temp_C)
```

```{r}
## WILL NEED TO EDIT THIS ONCE I HAVE DATA FROM SITE!
bs_trans_ub_aspects <- bs_trans_ub_alldata %>%
  mutate(TIMESTAMP = paste0(TIMESTAMP, ":00")) %>%  # Add missing seconds
  mutate(datetime = mdy_hms(TIMESTAMP, tz = "MST")) %>%
  select(-TIMESTAMP) %>%
  select(datetime, everything()) %>%
  pivot_longer(cols = matches("^UB"),  # Use matches() for multiple prefixes
  names_to = "tree_id",
  values_to = "temp_C") %>%
  #add in aspect column
  mutate(aspect = ifelse(
    tree_id %in% c(
      "UB_N_N_2_Avg",
      "UB_N_N_05_Avg"), 
    "north",
    "south"
  )) %>%
  # Add tree_name column based on the start of tree_id
  mutate(tree_name = "ub_n", burn_status = "live", zone = "TSZ") %>%
  mutate(sensor_height_m = case_when(
    str_ends(tree_id, "2_Avg") ~ 2,  # If tree_id ends with "1_Avg"
    str_ends(tree_id, "05_Avg") ~ 0.5,  # If tree_id ends with "05_Avg"
    TRUE ~ NA_real_  # If neither condition is met, set to NA
  )) %>%
  mutate(day = as.Date(datetime, tz = "MST")) %>%
  select(datetime, day, zone, burn_status, tree_name, aspect, sensor_height_m, temp_C)
```

\*\* add in phase next! Need to download p_solar_times then do a mutate to add phase to all of the above datasets -\> but make a new version labeled XXX_r?

Adding in phase

```{r}
# Specify PSZ and TSZ locations
##pers_b - from 105E location!
p_b_lat <- 40.56379284
p_b_lon <- -105.867215
##pers_ub - from 105E locations
p_ub_lat <- 40.5638075
p_ub_lon <- -105.8705715

##trans_b - from 105E locations!
t_b_lat <- 40.60600629
t_b_lon <- -105.8333335

##trans_ub - from 105E locations!
t_ub_lat <- 40.568396
t_ub_lon <- -105.587383

## use the p_b_lat/long since all of these dates/times are almost identical for sunrise/sunset

# Create a date sequence from February 2nd to May 24th, 2024 for solar exposure
solex_dates <- seq(as.Date("2024-10-01"), as.Date("2025-07-01"), by = "day")

# Get solar times for each day
p_b_solar_times_2025 <- getSunlightTimes(date = solex_dates, lat = p_b_lat, lon = p_b_lon, tz = "MST") %>%
  mutate(
    day = as.Date(date),  # Ensure 'day' is of Date type for joining
    sunrise = with_tz(as.POSIXct(sunrise), tzone = "MST"),  # Set MST timezone
    sunset = with_tz(as.POSIXct(sunset), tzone = "MST")     # Set MST timezone
  )
```

```{r}
#cutting the sunrise/sunset times to the current date range
#PSZ burned
p_b_solar_times_2025_V2 <- p_b_solar_times_2025 %>%
  mutate(day = as.Date(date)) %>%
  filter(day >= as.Date("2024-12-18") & day <= as.Date("2025-01-10")) %>% 
  mutate(day = as.Date(day, format = "%Y-%m-%d"))

#TSZ burned
t_b_solar_times_2025 <- p_b_solar_times_2025 %>%
  mutate(day = as.Date(date)) %>%
  filter(day >= as.Date("2024-11-13") & day <= as.Date("2025-01-15")) %>% 
  mutate(day = as.Date(day, format = "%Y-%m-%d"))

#PSZ unburned
p_ub_solar_times_2025 <- p_b_solar_times_2025 %>%
  mutate(day = as.Date(date)) %>%
  filter(day >= as.Date("2024-11-12") & day <= as.Date("2025-01-09")) %>% 
  mutate(day = as.Date(day, format = "%Y-%m-%d"))

#TSZ unburned - DONT HAVE DATA YET!
t_ub_solar_times_2025 <- p_b_solar_times_2025 %>%
  mutate(day = as.Date(date)) %>%
  filter(day >= as.Date("2024-11-12") & day <= as.Date("2025-01-09")) %>% 
  mutate(day = as.Date(day, format = "%Y-%m-%d"))
```

```{r}
#PSZ burned data with day/night
bs_pers_b_aspects_filtered <- bs_pers_b_aspects %>%
  mutate(day = as.Date(day, format = ("%Y-%m-%d"))) %>%
  inner_join(p_b_solar_times_2025_V2, by = "day") %>%
  #creating a phase column that parses by daily sunrise/sunset times
  mutate(
    phase = case_when(
      datetime >= sunrise & datetime < sunset ~ "day",
      datetime < sunrise ~ "night",
      datetime >= sunset ~ "night",
      TRUE ~ NA_character_)) %>%
  select(datetime, day, phase, zone, burn_status, tree_name, aspect, sensor_height_m, temp_C, solarNoon, sunrise, sunset, dawn, dusk)
```

```{r}
#TSZ burned data with day/night
bs_trans_b_aspects_filtered <- bs_trans_b_aspects %>%
  mutate(day = as.Date(day, format = ("%Y-%m-%d"))) %>%
  inner_join(t_b_solar_times_2025, by = "day") %>%
  #creating a phase column that parses by daily sunrise/sunset times
  mutate(
    phase = case_when(
      datetime >= sunrise & datetime < sunset ~ "day",
      datetime < sunrise ~ "night",
      datetime >= sunset ~ "night",
      TRUE ~ NA_character_)) %>%
  select(datetime, day, phase, zone, burn_status, tree_name, aspect, sensor_height_m, temp_C, solarNoon, sunrise, sunset, dawn, dusk)
```

```{r}
#PSZ unburned data with day/night
bs_pers_ub_aspects_filtered <- bs_pers_ub_aspects %>%
  mutate(day = as.Date(day, format = ("%Y-%m-%d"))) %>%
  inner_join(p_ub_solar_times_2025, by = "day") %>%
  #creating a phase column that parses by daily sunrise/sunset times
  mutate(
    phase = case_when(
      datetime >= sunrise & datetime < sunset ~ "day",
      datetime < sunrise ~ "night",
      datetime >= sunset ~ "night",
      TRUE ~ NA_character_)) %>%
  select(datetime, day, phase, zone, burn_status, tree_name, aspect, sensor_height_m, temp_C, solarNoon, sunrise, sunset, dawn, dusk)
```

Exporting filtered datasets

```{r}
#datasets with full extent as of now, but will need to rerun when I upload new data to the composite files
write_csv(bs_pers_b_aspects_filtered, "nsf/105E_pers_burned/105E_pers_b_bs_r.csv")

write_csv(bs_trans_b_aspects_filtered, "nsf/105E_trans_burned/105E_trans_b_bs_r.csv")

write_csv(bs_pers_ub_aspects_filtered, "nsf/105E_pers_unburned/105E_pers_ub_bs_r.csv")

write_csv(bs_trans_ub_aspects_filtered, "nsf/105E_trans_unburned/105E_trans_ub_bs_r.csv")
```
